### 2024-11-12 src/tex/algo.tex
[showyourwork] overleaf sync

\begin{algorithm}
\caption{Basic pseudo-code for Bayesian optimization}
\label{alg:BayesOpt}
\begin{algorithmic}[1] % [1] turns on line numbering
    \State Place Gaussian process $\mathcal{GP}$ prior on $\Lmain$
    \State Observe $\Lmain$ at $N_0$ points randomly sampled.
    \While{$n \leq N$}
        \State Update the posterior probability distribution on $\Lmain$ using all available data
        \State Let $\lambda_{\text{best}}$ be a maximizer of the acquisition function over $\lambda$, where the acquisition function is computed using the current posterior distribution.
        \State Observe $L_{\text{best}} = \mathcal{L}(\mathcal{D}|\lambda_{\text{best}})$.
        \State Increment $n$
    \EndWhile
    \State Return a solution: $\Lsurr \sim \mathcal{GP}(\lambda)$
\end{algorithmic}
\end{algorithm}